---
title: "Participation 1-30-2020"
author: "Kaitlin"
date: "1/30/2020"
output: html_document
---

Installation and Set Up

```{r}
#install.packages("tidyverse")
#install.packages("scales")
#install.packages("tsibble")

suppressPackageStartupMessages(library(tidyverse))
suppressPackageStartupMessages(library(gapminder))
suppressPackageStartupMessages(library(scales))
knitr::opts_chunk$set(fig.width = 5, fig.height = 4, fig.align = "center")
```

Code Example

```{r}
ggplot(gapminder, aes(gdpPercap, lifeExp)) +
  geom_point(alpha = 0.1) +
  scale_x_log10("GDP per capita", labels = scales::dollar_format()) +
  theme_bw() +
  ylab("Life Expectancy")
```


In class Practice - Naming Components

```{r, fig.width=5, fig.height=2}
gapminder %>% 
  filter(year==2007) %>% 
  mutate(continent=fct_infreq(continent)) %>% 
  ggplot(aes(continent)) +
  geom_bar() +
  theme_bw()
```

With the output plot
| Grammar Component     | Specification |
|-----------------------|---------------|
| __data__              | `gapminder`  |
| __aesthetic mapping__ |              |
| __geometric object__  | Bar          |
| scale                 | Linear       |
| statistical transform | none         |
| coordinate system     | rectangular  |
| facetting             | none         |

~~~~~~~~~~~~~~~~~~~~~~~~~~~~~

Class Walkthrough of a Making Scatter Plot- see notes

Step 1 - The Basics

```{r}
ggplot(gapminder, aes(gdpPercap, lifeExp))
```

Step 2 - geom_point

```{r}
ggplot(gapminder, aes(gdpPercap, lifeExp)) +
  geom_point()
```

Step 3 - Alpha and Color

```{r}
ggplot(gapminder, aes(gdpPercap, lifeExp)) +
  geom_point(alpha =.1, aes(color = continent))

```

Step 4 - Scale 

```{r}
ggplot(gapminder, aes(gdpPercap, lifeExp)) +
  geom_point(alpha =.1, aes(color = continent)) +
  scale_x_log10()
```

using log 10 transformation - not sure what that math is..? It's fine.

Step 5 - Specifying Scale and Label

```{r}
ggplot(gapminder, aes(gdpPercap, lifeExp)) +
  geom_point(alpha =.1, aes(color = continent)) +
  scale_x_log10("GDP per capita", labels=scales::dollar_format())
```
  

Okay Cool! We're happy with this! Now... Lets make it prettier.

Step 6 - Theme & Adding $ to GDP
```{r}
ggplot(gapminder, aes(gdpPercap, lifeExp)) +
  geom_point(alpha =.1, aes(color=continent)) +
  scale_x_log10("GDP Per Capita", labels=scales::dollar_format()) +
  theme_grey()
```

Step 7 - Naming Y Axis

```{r}
ggplot(gapminder, aes(gdpPercap, lifeExp)) +
  geom_point(alpha =.1, aes(color=continent)) +
  scale_x_log10("GDP Per Capita", labels=scales::dollar_format()) +
  theme_grey() +
  ylab("Life Expectancy")
```


Step 8 - Figure Size
```{r, fig.width=7, fig.height=5}
ggplot(gapminder, aes(gdpPercap, lifeExp)) +
  geom_point(alpha =.1, aes(color=continent)) +
  scale_x_log10("GDP Per Capita", labels=scales::dollar_format()) +
  theme_grey() +
  ylab("Life Expectancy")
```

