---
title: 's05: Some More `dplyr` Exercise'
author: "Kaitlin White"
date: "2/13/2020"
output:
  html_document:
    keep_md: true
    theme: paper
---

# Participation 2-13-2020


## Setting Up

### Error Allowance

<!---The following chunk allows errors when knitting--->
```{r allow errors, echo = FALSE}
knitr::opts_chunk$set(error = TRUE)
```

### Reminders
**When you make an Rmd file for participation or homework, be sure to do this**:

1. Change the file output to both html and md _documents_ (not notebook).
  - See the `keep_md: TRUE` argument above.

2. `knit` the document. 

3. Stage and commit the Rmd and knitted documents.

### Packages
```{r load packages, warning = FALSE, message = FALSE}
library(tidyverse)
library(gapminder)
# install.packages("tibble")
```
    

## Worksheet

### 1.

`select()`, `rename()`, `filter()`, `mutate()`, and a little plotting

Let's use the `mtcars` dataset. Complete the following tasks. Chain together
all of the commands in a task using the pipe `%>%`. Show the miles per gallon and horsepower for cars with 6 cylinders. Also convert the data frame to a tibble (keep the rownames and store them in the tibble with a descriptive variable name). Store this result as a new object with a descriptive object name.

```{r}
mtcars
```

```{r}
mtcars_2 <- mtcars %>%
  as_tibble(rownames = "model") %>%
  filter(cyl == 6) %>%
  select(model, hp, mpg) %>% 
  rename("Model"="model","Horsepower"="hp", "MPG"="mpg") %>%
  print()
```

### 2. 

Print the results from Task 1 in an appealing way by using `knitr::kable()`.

```{r}
knitr::kable(mtcars_2)
```

Let's use the `iris` dataset. Complete the following tasks. Chain together all of the commands in a task using the pipe `%>%`.

```{r}
iris
```


### 3.

Rename the variables to be all lowercase and to separate words with "_" instead of ".". Put the species name variable first. Store this result as a new object.

```{r}
iris_tidy <- iris %>%
  select("species"="Species",
         "sepal_length"="Sepal.Length", 
         "sepal_width"="Sepal.Width",
         "petal_length"="Petal.Length",
         "petal_width"="Petal.Width") %>%
  print()
  
```


### 4. 

Using the data from Task 3, plot the sepal width for each species. Perhaps use a boxplot or a jitter plot (or both overlaid!). Be sure to format the axis labels nicely.

```{r}
iris_tidy %>%
  ggplot(aes(x = sepal_width, y = species)) +
    geom_jitter() 
```


### 5. 

`iris` expresses all of the measurements in centimeters. Convert them to inches (1 in = 2.54 cm). Store this dataset as a new object.

```{r}
iris_tidy_inches <- iris_tidy %>%
  mutate(sepal_length = sepal_length/2.54,
         sepal_width = sepal_width/2.54,
         petal_length = petal_length/2.54,
         petal_width = petal_width/2.54,) %>%
  print()
  
```


### 6.

Using the data from Task 5, plot the relationship between sepal width and sepal length. Indicate species using color and point shape.

```{r}
iris_tidy_inches %>%
  ggplot(aes(x = sepal_width, y = sepal_length)) +
    geom_point(aes(color = species, shape = species))
```


### 7.

Using the data from Task 5, plot the relationship between sepal width and sepal length. This time, separate each species into a different subplot (facet).

```{r}
iris_tidy_inches %>%
  ggplot(aes(x = sepal_width, y = sepal_length)) +
    geom_point(aes(color = species, shape = species)) +
  facet_grid(rows = vars(species))
```


<br><br>

## Back to Guide Again

Let's head back to the guide at the section on `summarize()`.

<br><br>


## Exercises for grouped data frames

Let's do some practice with grouping (and ungrouping) and summarizing data frames!


### 1.

   (a) What's the minimum life expectancy for each continent and each year? 
   (b) Add the corresponding country to the tibble, too. 
   (c) Arrange by min life expectancy.

```{r}
gapminder %>% 
  group_by(continent, year) %>% 
  summarize(min_life = min(lifeExp)) %>%
  arrange(min_life) %>%
  transmute(min_life_decade = round(min_life, -1))
```


### 2.

Let's compute the mean Agreeableness score across items for each participant  in the `psych::bfi` dataset. Be sure to handle `NA`!

```{r}
psych::bfi
```

```{r}
psych::bfi %>%
  as_tibble() %>%
  select(A1:A5) %>%
  rowwise() %>%
  mutate(A_Mean = mean(c(A1, A2, A3, A4, A5), na.rm = TRUE))
```

Now compute mean scores for Conscientiousness, as well as `sd` and `min` scores for reach person.

```{r}
psych::bfi %>%
  as_tibble() %>%
  select(C1:C5) %>%
  rowwise() %>%
  mutate(C_Mean = mean(c(C1, C2, C3, C4, C5), na.rm = TRUE), 
         C_SD = sd(c(C1, C2, C3, C4, C5), na.rm = TRUE), 
         C_Min = min(c(C1, C2, C3, C4, C5), na.rm = TRUE))
```

Some functions are **vectorized**, so you don't need `rowwise()`. 
For example, `pmin()` computes the "parallel min" across the vectors it receives:

```{r}
psych::bfi %>%
  as_tibble() %>%
  select(C1:C5) %>%
  pmin()
```

**There are a few other ways to do this sort of computation.**

`rowMeans()` computes the mean of each row of a data frame. We can use it by
putting `select()` inside of `mutate()`:


```{r}

```

**In the development version of `dplyr`, there are some functions to make**
**this approach easier.**

```
remotes::install_github("tidyverse/dplyr")
```

```{r, eval = FALSE}

```

### 3.
Let's use `psych::bfi` and make a new data frame that has (1) each participant's educational level (convert it to a categorical variable using `factor*()`) and the mean score for each of the Big Five scales for each participant. Store this data frame as a new object.
   
```{r}
```

###4. 
Use the data from Task 3 to summarize the distributions of Big Five scores for each educational level (e.g., report the mean, sd, min, and max for each score in each group). Also report the sample size within each group.
   
```{r}

```  
