---
title: "s06 Exercises: Pivoting and tidy data"
author: "Kaitlin White"
date: "2-20-2020"
output: 
  html_document:
    keep_md: true
    theme: paper
---
## Set up
```{r, warning = FALSE, message = FALSE}
library(tidyverse)
lotr  <- read_csv("https://raw.githubusercontent.com/jennybc/lotr-tidy/master/data/lotr_tidy.csv")
guest <- read_csv("https://raw.githubusercontent.com/USF-Psych-DataSci/Classroom/master/data/wedding/attend.csv")
email <- read_csv("https://raw.githubusercontent.com/USF-Psych-DataSci/Classroom/master/data/wedding/emails.csv")
```

<!---The following chunk allows errors when knitting--->

```{r allow errors, echo = FALSE}
knitr::opts_chunk$set(error = TRUE, warning = FALSE)
```


## Exercise 1: Univariate Pivoting

Consider the Lord of the Rings data:

```{r}
lotr
```

### 1. Would you say this data is in tidy format?
   
   Yes, this data is tidy. The requirement to be "tidy" data are as follows
   -1 Each row is an observation. 
   -2 Each column is a variable.
   -3 Each cell is a value.
   
   1) In this table, each row is talking about a specific observation. IN X movie, this gender of this race spoke ____ amount of words. 
   2) Each column only contains one variable.
   3) Each cell holds a specific value that is relevant to the observation.

### 2. Widen the data so that we see the words spoken by each race, by putting race as its own column.

```{r}
(lotr_wide <- lotr %>% 
  pivot_wider(id_cols = c(-Race, -Words), 
              names_from = Race, 
              values_from = Words))
```

### 3. Re-lengthen the wide LOTR data from Question 2 above.

```{r}
lotr_wide_lengthened <-
  lotr_wide %>% 
  pivot_longer(cols = c(Elf, Hobbit, Man),
               names_to  = "Race",
               values_to = "Words") %>% 
print()
```

## Exercise 2: Multivariate Pivoting

Congratulations, you're getting married! In addition to the wedding, you've decided to hold two other events: a day-of brunch and a day-before round of golf. You've made a guestlist of attendance so far, along with food preference for the food events (wedding and brunch).

```{r}
guest %>% 
  DT::datatable(rownames = FALSE)
```

1. Put "meal" and "attendance" as their own columns, with the events living in a new column.

```{r}
(guest_long <- 
guest %>% 
  pivot_longer(cols = c(-party, -name), 
               names_to = c(".value", "Event"), 
               names_sep = "_"))
```

2. Use `tidyr::separate()` to split the name into two columns: "first" and 
"last". Then, re-unite them with `tidyr::unite()`.

```{r}
guest_long %>% 
  separate(name, into = c("first name", "last name"), sep = " ", remove = TRUE)
```


```{r}
guest_long %>% 
  separate(name, into = c("first name", "last name"), sep = " ", remove = TRUE) %>% 
  unite(col = "name", c("first name", "last name"), sep = " ", remove = TRUE)
```


3. Which parties still have a "PENDING" status for all members and all events?

```{r}
guest_long %>% 
  group_by(party) %>% 
  summarize(all_pending = all(attendance == "PENDING"))
```

4. Which parties still have a "PENDING" status for all members for the wedding?

```{r}
guest %>% 
  group_by(party) %>% 
  summarize(pending_wedding = all(attendance_wedding == "PENDING"))
```


5. Put the data back to the way it was.

```{r}
guest_long %>% 
  pivot_wider(id_cols = c(party, name), 
              names_from = Event, 
              names_sep = "_", 
              values_from = c(attendance, meal))
```

6. You also have a list of emails for each party, in this worksheet under the  object `email`. Change this so that each person gets their own row. 
   Use `tidyr::separate_rows()`

```{r}
email
```

```{r}
email %>% 
  separate_rows(guest, sep = ",")
```

