---
title: "Notes 2/20/2020"
author: "Kaitlin"
date: "2/20/2020"
output: 
  html_document:
    keed_md: true
    theme: paper       
---

## Set Up
```{r}
library(tidyverse)
```
~~~~~~~

# Tidy Data 

Data is tidy if:
1 Each row is an observation.
2 Each column is a variable.
3 Each cell is a value.

This means that each value belongs to exactly one variable and one observation.


```{r}
datasets::HairEyeColor
```

~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~


# Pivoting

```{r}
HairEyeColor_untidy <- HairEyeColor %>% 
  mutate(eye = str_c(eye, "_eyed")) %>% 
  pivot_wider(id_cols = hair, names_from = eye, values_from = n)
```

To fix this, we need to convert the eye colour columns to two columns:

one column to hold the eye colour (column names),
one column to hold the values.

~~~

## pivot_longer() 
takes a data frame, and returns a data frame. The arguments after the data argument that we’ll need are:

**cols** for the column names that we want to turn into a single column.
**names_to:** the old column names are going to a new column. What should this new column be named? (optional, but highly recommended)
**values_to:** the values underneath the old columns are going to a new column. What should this new column be named? (optional, but highly recommended)

```{r}
HairEyeColor_untidy %>% 
  pivot_longer(cols = c(Blue_eyed, Brown_eyed, Green_eyed, Hazel_eyed),
               names_to  = "eye",
               values_to = "n")
```


Better is to use helper functions from the tidyselect package. In this case, we know the columns contain the text “eyed”, so let’s use tidyselect::contains():

```{r}
haireye_untidy %>% 
  pivot_longer(cols      = contains("eyed"),
               names_to  = "eye",
               values_to = "n")
```

Yet another way is to indicate everything except the hair column:

```{r}
haireye_untidy %>% 
  pivot_longer(cols      = -hair,
               names_to  = "eye",
               values_to = "n")
```               
~~~
    
## pivot_wider()
Like pivot_longer(), pivot_wider() takes a data frame and returns a data frame. The arguments after the data argument that we’ll need are:

id_cols: The columns you would like to keep. If widening to make data tidy, then this is an identifier for an observation.
names_from: the new column names are coming from an old column. Which column is this?
values_from: the values under the new columns are coming from an old column. Which column is this?

```{r}
HairEyeColor %>% 
  pivot_wider(id_cols = hair, 
              names_from = eye, 
              values_from = n)    
```


~~~

# Multivariate Pivoting - Pivoting Mutliple Variables

Now let’s consider the case when more than one variable are “out of place” – perhaps there are multiple variables per row, and/or multiple observations per row.

## pivot_longer() MV

To obtain the case where two (or more) variables are contained in column names, here’s how we specify the arguments of pivot_longer():

cols: As usual.
names_sep: What is separating the variables in the column names?
names_to: The old columns are going to be put into new columns, after being separated. What should those columns be named?
values_to: As usual.

```{r}
iris2 %>% 
  pivot_longer(cols      = c(-species, -id), 
               names_to  = c("part", "dimension"),
               names_sep = "_",
               values_to = "measurement")
```


## Pivot_wider() MV
To obtain the case where multiple observations are contained in one row, here’s how to specify the arguments of pivot_longer():

cols: As usual.
names_sep: As above.
names_to: As above, except this time, one part of the old column names are going to stay as columns (in this case, “length” and “width”). Indicate ".value" instead of a new column name.
values_to: Not needed! You’ve already indicated that using the ".value" placeholder.

iris2 %>% 
  pivot_longer(cols      = c(-id, -species), 
               names_to  = c("part", ".value"), 
               names_sep = "_")