---
title: "Activity 1"
author: "Kaitlin"
date: "1/30/2020"
output: html_document
---
# Participation Activities Jan 30th 2020

## Installation and Set Up
```{r}
#install.packages("tidyverse")
#install.packages("scales")
#install.packages("tsibble")

suppressPackageStartupMessages(library(tidyverse))
suppressPackageStartupMessages(library(gapminder))
suppressPackageStartupMessages(library(scales))
knitr::opts_chunk$set(fig.width = 5, fig.height = 4, fig.align = "center")
```

## Activities!

### 2(a) - Write Your Own!

 Specifically, the plot should have the following grammar components:

| Grammar Component     | Specification |
|-----------------------|---------------|
| __data__              | `mauna` |
| __aesthetic mapping__ | x: month, y: conc |
| __geometric object__  | lines |
| scale                 | linear |
| statistical transform | none |
| coordinate system     | rectangular |
| facetting             | none |

#### Opening Mauna - 

```{r}
(mauna <- tsibble::as_tsibble(co2) %>% 
   rename(month = index, conc = value))
```

#### Making the plot - 

```{r, fig.width= 7 , fig.height= 5}
ggplot (mauna, aes(month, conc)) +
  geom_line() +
  theme_classic() 
```


### 2(b)- Fix The Code

Code Provided:
{r, fig.with = 5, fig.height = 2}
ggplot(mauna, aes(y = month)) +
  geom_line(aes(FILL_THIS_IN))


Fixed:
```{r, fig.with = 5, fig.height = 2}
ggplot(mauna, aes(month, conc)) +
  geom_line()
```



### 2(c) - Storing & P

You can store the output of the plot in a variable, too. Store the plot from 2(a) in the variable named `p`, then add a layer to `p` that adds green points to the plot.

```{r, fig.width = 5, fig.height = 2}
p<-ggplot (mauna, aes(month, conc)) +
  geom_line() +
  theme_classic() 
p +
  geom_point(color="green")
```

### 2(d)

What's wrong with the following code? Fix it.

Code Provided:
{r, fig.width = 5, fig.height = 2}
ggplot(gapminder) +
  geom_point(x = gdpPercap, y = lifeExp, alpha = 0.1)

Fixed:
```{r, fig.width = 5, fig.height = 2}
ggplot(gapminder, aes(gdpPercap, lifeExp))+
  geom_point(alpha = 0.1)
```


### 2(e) BONUS

So you're a ggplot2 pro? Then, let's see this plot adapted to polar coordinates. Specifically:

- angle is month (January through December)
- radius is CO$_2$ concentration

The plot should look like a spiral, or concentric circles. 

```{r, fig.width = 5, fig.height = 2}
ggplot (mauna, aes(month, conc)) +
  geom_????(angle=month.abb, radius=co2) +
  theme_classic() 
```
