---
title: R Markdown Exploration
author: Kaitlin
date: last updated "`r format(Sys.time(), '%d %B, %Y')`"
output:
  html_document:
    keep_md: true
    theme: cerulean
    toc: true
    toc_float: true
---
This document is from s02 on January 23rd 2020 - this day covering an introduction to rmarkdown. 


This is a document. Words words words. Yay. Words.


This is a code block - __r__ is the reference to what programming language is in use. The lower commands are a section of R code. I'm using this chunk to install packages.

```{r} 
# install.packages("rmarkdown")
# install.packages
# 
# install.packages('gapminder')
# install.packages('DT')
# install.packages('tibble')
```


Now ill use the library function to get ready to use them.
```{r}
library(gapminder)
library(DT)
library(tibble)
```


```{r}
gapminder
````

using mtcars data set - different ways to look at mpg - by graph, by numbers, and avg MPG formula
```{r}
mtcars

#mtcars["mpg"]
#mtcars[["mpg"]]
#mtcars$mpg

mean(mtcars[["mpg"]])
```
 
 
 Using tibble - not sure what this did?
```{r}
tibble::as_tibble(mtcars)
```


Did some editing to YAML